\textbf{Estimating the proportion of beneficial mutations that are not adaptive in mammals}

### 2024-12-16 manuscript/S2 File.tex
Split Supplementary materials into 4 files.

\begin{document}
    \maketitle
    \part*{Supplementary materials}
    \tableofcontents

    \newpage
    \listoffigures

    \newpage
    \listoftables

    \clearpage
    \section{Mutation-selection codon models}\label{sec:mutsel}

    \subsection{Prior distributions and parameters of the model}\label{subsec:priors}

    We here present the Mutation-selection codon model as a Bayesian hierarchical model, including the prior distributions and the parameters of the model.
    This hierarchical model is formally represented as directed acyclic graph of dependencies between variables, depicted in Figure~\ref{mutsel-illustration}:

        \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=0.6cm and 1.2cm,semithick]
            \tikzstyle{every state}=[]

            \node[state] (P) {$\Probmatrix\branchsiteexp$};
            \node[state] (Q) [below right=of P] {$\Submatrix\branchsiteexp$};
            \node[state] (R) [below right=of Q] {$\Mutmatrix$};
            \node[state] (BL) [BLUE, right=of P] {$\branchlength\branchexp$};
            \node[res] (BLH) [RED, right=of BL] {$ \lambda, \tau $};
            \node[state] (f) [right=of Q] {$\Fit\siteexp $};
            \node[state] (Ex) [BLUE, below right=of R] {$\Exchan$};
            \node[state] (Equi) [BLUE, right=of R] {$\Mutequi$};
            \node[state] (Base) [BLUE, above right=of f] {$\Profile\catexp$};
            \node[state] (cat) [BLUE, right=of f] {$\catsite$};
            \node[res] (ExH) [RED, right=of Ex] {$1,1,1,1,1,1$};
            \node[res] (EquiH) [RED, right=of Equi] {$1,1,1,1$};
            \node[state] (baseH) [RED, right=of Base] {$\concentrationProfile, \centerProfile $};
            \node[state] (sb) [BLUE, right=of cat] {$\StickBreaking$};
            \node[state] (sbH) [RED, right=of sb] {$\stickbreakinghyper$};

            \path
            (Q) edge [black] node [above right] {\ref{eq:Probmatrix}} (P)
            (BL) edge [black] node [] {} (P)
            (R) edge [black] node {} (Q)
            (f) edge [black] node [above] {\ref{eq:subrates}} (Q)
            (Ex) edge [black] node [] {} (R)
            (Equi) edge [black] node [below] {\ref{eq:gtr-mutrates}} (R)
            (Base) edge [black] node {} (f)
            (cat) edge [black] node [above] {\ref{eq:sitefitness}} (f)
            (BLH) edge [dashed, BLUE] node [above] {\ref{eq:branchlength}} (BL)
            (ExH) edge [dashed, BLUE] node [above] {\ref{eq:DistribExchan}} (Ex)
            (EquiH) edge [dashed, BLUE] node [above] {\ref{eq:DistribMutequi}} (Equi)
            (baseH) edge [dashed, BLUE] node [above] {\ref{eq:DistribBase}} (Base)
            (sb) edge [dashed, BLUE] node [above] {\ref{eq:DistribMultinomial}} (cat)
            (sbH) edge [dashed, BLUE] node [above] {\ref{eq:DistribStickBreaking},\ref{eq:Beta}} (sb);
        \end{tikzpicture}
        \captionof{figure}[Bayesian hierarchical model.]{
    \textbf{Bayesian hierarchical model.} Nodes of the directed acyclic graph are the variables, and edges are the functions.
    Hyper-parameters are depicted in {\color{RED}{red}} circles, random variables in {\color{BLUE}{blue}} circles, and transformed variables in black.
    {\color{BLUE}{Blue}} dashed line denotes a drawing from a random distribution, and black solid lines denote a function.
    All the nodes pointing toward a given node (upstream) are its dependencies which determine its distribution.
    The other way around, following the arrows in the DAG (downstream), simple {prior} distributions are combined together to form more complex joint {prior} distributions which ultimately define the {prior} distribution of the model.
    Equations~\ref{eq:gtr-mutrates} to~\ref{eq:Probmatrix} describe the relationship between variables (sections~\ref{sec:nuc} to~\ref{sec:codon}).
    \label{mutsel-illustration}}
    \end{center}

    \subsection{Nucleotide mutation rates}
    \label{sec:nuc}
    The generalized time-reversible (GTR) nucleotide mutation rate matrix $\Mutmatrix$ is a function of the nucleotide frequencies $\Mutequi$ and the symmetric exchangeability rates $\Exchan$~\cite{tavare_probabilistic_1986}.
    $\Mutequi = (\mutequi_A , \mutequi_C , \mutequi_G , \mutequi_T)$ is the equilibrium base frequency vector, giving the frequency at which each base occurs at each site.
    $\Exchan = \left( \exchan_{AC}, \exchan_{AG}, \exchan_{AT}, \exchan_{CG}, \exchan_{CT}, \exchan_{GT}\right)$ is the vector of exchangeabilities between nucleotides.
    Altogether, the rate matrix is:
    \begin{equation}
        \label{eq:gtr-mutrates}
        \Mutmatrix =
        \begin{blockarray}{ccccc}
            & A & C & G & T \\
            \begin{block}{c(cccc)}
                A & -\mu_A & {\exchan_{AC}\mutequi_C} & {\exchan_{AG}\mutequi_G} & {\exchan_{AT}\mutequi_T} \\
                C & {\exchan_{AC}\mutequi_A} &                        -\mu_C & {\exchan_{CG}\mutequi_G} & {\exchan_{CT}\mutequi_T} \\
                G & {\exchan_{AG}\mutequi_A} & {\exchan_{CG}\mutequi_C} &                        -\mu_G & {\exchan_{GT}\mutequi_T} \\
                T & {\exchan_{AT}\mutequi_A} & {\exchan_{CT}\mutequi_C} & {\exchan_{GT}\mutequi_G} & -\mu_T\\
            \end{block}
        \end{blockarray}
    \end{equation}
    By definition, the sum of the entries in each row of the nucleotide rate matrix $\Mutmatrix$ is equal to $0$, giving the diagonal entries:
    \begin{equation}
        \mu_a = \sum\limits_{ b \neq a, b \in \{A, C, G, T\} } \mutmatrix_{a,b}
    \end{equation}
    The {prior} on the exchangeabilities $\Exchan$ is a uniform Dirichlet distribution of dimension $6$:
    \begin{equation}
        \label{eq:DistribExchan}
        \Exchan \sim \text{Dir}\left( 1,1,1,1,1,1 \right).
    \end{equation}
    The {prior} on the equilibrium base frequencies $\Mutequi$ is a uniform Dirichlet distribution of dimension $4$:
    \begin{equation}
        \label{eq:DistribMutequi}
        \Mutequi \sim \text{Dir}\left( 1,1,1,1 \right)
    \end{equation}
    The general time-reversible nucleotide matrix is normalized with a total flow of $1$:
    \begin{equation}
        \sum\limits_{a \in \{A, C, G, T\}} - \mutequi_a \mutmatrix_{a,a} = 1,
    \end{equation}
    such that we expect $1$ {substitution} per unit of branch length.

    \subsection{Site-specific amino-acid fitness profiles}
    \label{sec:profiles}
    Site-specific amino-acid fitness profiles are assumed i.i.d. from a mixture model, itself endowed with a truncated Dirichlet process prior.
    Specifically, the mixture has $\Ncat$ components ($\Ncat = 30$ by default).
    The prior on component weights~($\StickBreaking$) is modeled using a stick-breaking process, truncated at $\Ncat$ and of parameter $\stickbreakinghyper=1$:
    \begin{align}
        \label{eq:DistribStickBreaking}
        \begin{split}
            & \StickBreaking \sim \text{StickBreaking}\left( \Ncat, \stickbreakinghyper \right)\\
            \iff & \stickbreaking_{\cat} = \stick_{\cat}\cdot \prod _{{\indice=1}}^{{\cat-1}}\left(1-\stick_{\indice}\right),\ \Setcat,
        \end{split}
    \end{align}
    where $\stick_{\cat}$ are i.i.d. from a beta distribution
    \begin{equation}
        \label{eq:Beta}
        \stick_{\cat} \sim \text{Beta}\left( 1, \stickbreakinghyper \right),\ \Setcat.
    \end{equation}
    Of note, the weights decrease geometrically in expectation, at rate $\stickbreakinghyper$, such that lower values of $\stickbreakinghyper$ induce more heterogeneous distributions of weights.

    Each component of the mixture defines a 20-dimensional fitness profile $\Profile\catexp$ (summing to $1$), for $ \Setcat$.
    These fitness profiles are i.i.d. from a Dirichlet of center $\centerProfile=1$ and concentration $\concentrationProfile=1$:
    \begin{equation}
        \label{eq:DistribBase}
        \Profile\catexp \sim \text{Dir}\left( \centerProfile,\ \concentrationProfile \right),\ \Setcat.
    \end{equation}

    Site allocations to the mixture components $\catsite \in \catInterval $, for $\Setsite$ running over the $\Nsite$ sites of the alignment, are i.i.d. multinomial of parameter $\StickBreaking$:
    \begin{align}
        \label{eq:DistribMultinomial}
        \catMultiVar \sim \text{Multinomial}\left( \StickBreaking \right), \\
        \text{ where } \catmultivar_{\cat} = \sum_{\Setsite} \mathbb{1}_{\catsite = \cat}
    \end{align}

    For a given parameter configuration for the mixture, the scaled fitness $\Fit\siteexp$ at site $\site$, are obtained by taking the logarithm of fitness assigned to this site:
    \begin{equation}
        \label{eq:sitefitness}
        \Fit\siteexp = \ln \left( \Profile^{\left( \catsite \right)} \right),\ \Setsite.
    \end{equation}

    \subsection{Branch length}
    The topology of the rooted phylogenetic tree is supposed to be known and is not estimated by the model.
    The branch lengths $\branchlength\branchexp$ are defined as the expected number of {neutral} substitutions per {DNA} site along a branch, each from a Gamma distribution of mean $\lambda=0.1$ and scale $\tau=1$:
    \begin{equation}
        \label{eq:branchlength}
        \branchlength\branchexp \sim \text{Gamma}\left( \lambda, \tau \right).
    \end{equation}

    \subsection{Codon substitution rates}
    \label{sec:codon}
    The mutation rate between codons $\ci$ and $\cj$, denoted $\mu_{\itoj}$ depends on the underlying nucleotide change between the codons.
    First, if codons $\ci$ and $\cj$ are not nearest-neighbours, $\mu_{\itoj}$ is equal to $0$.
    Second, if codons $\ci$ and $\cj$ are only one mutation away, $\mu_{\itoj}$ is given by the underlying nucleotide relative rate~(${\mutmatrix_{\itoj}}$).

    For a given site $\site$, the {codon} {substitution} rate matrix $\Submatrix\siteexp$ is given by:
    \begin{equation}
        \label{eq:subrates}
        \begin{dcases}
            \submatrix\siteexp_{\itoj} = 0\text{ if codons $\ci$ and $\cj$ are not nearest-neighbors,} \\
            \submatrix\siteexp_{\itoj} = \mu_{\itoj}\text{ if codons $\ci$ and $\cj$ are {synonymous},} \\
            \submatrix\siteexp_{\itoj} = \mu_{\itoj} \dfrac{ \fitj\siteexp - \fiti\siteexp}{{1 - \e^{\fiti\siteexp - \fitj\siteexp} }} \text{ if codons $\ci$ and $\cj$ are non-synonymous,}\\
            \submatrix\siteexp_{\ci, \ci} = - \sum\limits_{ \cj \neq \ci, \cj=1}^{61} \submatrix\siteexp_{\itoj}.
        \end{dcases}
    \end{equation}
    Together, the probability of transition between codons for a given branch $\branch$ and site $\site$ is:
    \begin{equation}
        \label{eq:Probmatrix}
        \Probmatrix\branchsiteexp = \e^{\branchlength\branchexp \Submatrix\siteexp},
    \end{equation}
    which are the matrices necessary to compute the {likelihood} of the data ($\data$) given the parameters of the model using the pruning algorithm.

    \subsection{Bayesian implementation}
    \label{sec:Bayesian}
    Bayesian inference was conducted using Markov Chain Monte Carlo (MCMC).
    Most phylogenetic {MCMC} samplers target the distribution over the model parameters given the sequence alignment, which means that they have to repeatedly invoke the pruning algorithm to recalculate the {likelihood} which is most often the limiting step of the {MCMC}.
    An alternative, which is used here, is to do the {MCMC} conditionally on the detailed {substitution} history $\subhistory$, thus doing the {MCMC} over the augmented configuration~($\subhistory$, $\data$), under the target distribution obtained by combining the mapping-based {likelihood} with the {prior} over model parameters.
    The key idea that makes this strategy efficient is that the mapping-based {likelihood} depends on compact summary statistics of $\subhistory$, leading to a very fast evaluation of the {likelihood}.
    On the other hand, this requires to implement more complex {MCMC} procedures that have to alternate between:
    \begin{enumerate}
        \item sampling $\subhistory$ conditionally on the data and the current parameter configuration.
        \item re-sampling the parameters conditionally on $\subhistory$.
    \end{enumerate}

    To implement the mapping-based {MCMC} sampling strategy, we first sample the detailed {substitution} history $\subhistory$ for all sites along the tree.
    Several methods exist for doing this~\cite{nielsen_mapping_2002,rodrigue_uniformization_2008}, which are used here in combination: first trying the accept-reject method of \textcite{nielsen_mapping_2002}, then switching to the uniformization approach of \textcite{rodrigue_uniformization_2008} if the first round has failed.

    Then, we write down the probability of $\subhistory$ given the parameters, and finally, we collect all factors that depend on some parameter of interest and make some simplifications.
    This ultimately leads to relatively compact sufficient statistics allowing for fast numerical evaluation of the likelihood~\cite{irvahn_phylogenetic_2014,davydov_state_2017}.

    \subsection{BayesCode software}\label{subsec:bayescode}
    In \textit{BayesCode} (\href{https://github.com/ThibaultLatrille/BayesCode}{github.com/ThibaultLatrille/BayesCode}, v1.3.1), we ran the mutation-selection codon models \textit{mutselomega} for 2000 points of MCMC with the options:
    \begin{scriptsize}
        \begin{verbatim}
 mutselomega ---omegashift 0.0 --ncat 30 -a my_alignment.phy -t my_tree.newick -u 2000 my_genename
        \end{verbatim}
    \end{scriptsize}
    The collection of site-specific fitness profiles ($\UniDimArray{F^{(i)}}, \forall i$) are then obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --ss my_genename
        \end{verbatim}
    \end{scriptsize}
    The gene-specific $4 \times 4$ nucleotide mutation rate matrix ($\UniDimArray{\mu}$) is also obtained by running \textit{readmutselomega}, reading 1000 points of MCMC (first 1000 are considered as burn-in) with the options:
    \begin{scriptsize}
        \begin{verbatim}
 readmutselomega --every 1 --until 2000 --burnin 1000 --nuc my_genename
        \end{verbatim}
    \end{scriptsize}

    \section{Beneficial mutations in the terminal lineages and populations}\label{sec:beneficial-mutations}

@@

    \section{Contrasting selection at the phylogenetic and population-genetic scales}

    \subsection{Selection at the population-genetic scale}
    The proportion of deleterious ($\ProbaPopDel$), nearly-neutral ($\ProbaPopNeu$) and of beneficial ($\ProbaPopBen$) mutations estimated at the population-genetic scale across the different populations is shown for each class of selection:
    $\SphyDel$ (Figure~\ref{fig:pop-neg}), $\SphyNeu$ (Figure~\ref{fig:pop-weak}) and $\SphyBen$ (Figure~\ref{fig:pop-pos}).

    \begin{center}
        \begin{minipage}{0.75\linewidth}
            \begin{minipage}{0.9\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/Theta.neg.stacked.pdf}
            \end{minipage}
            \begin{minipage}{0.09\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
            \end{minipage}
        \end{minipage}
    \captionof{figure}[Estimation of selection at the population scale for $\SphyDel$ mutations.]{\textbf{Estimation of selection at the population scale for $\bm{\SphyDel}$ mutations.} Each stack bar corresponds to the DFE of new mutations predicted to be $\SphyDel$  by the mutation selection model, in a given population. The dashed area represents the proportion of deleterious mutations ($\SpopDel$), the gray area that of neutral mutations ($\SpopNeu$), and the yellow area that of beneficial mutations ($\SpopBen$).\label{fig:pop-neg}}
    \end{center}

    \begin{center}
        \begin{minipage}{0.75\linewidth}
            \begin{minipage}{0.9\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/Theta.weak.stacked.pdf}
            \end{minipage}
            \begin{minipage}{0.09\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
            \end{minipage}
        \end{minipage}
    \captionof{figure}[Estimation of selection at the population scale for $\SphyNeu$ mutations.]{\textbf{Estimation of selection at the population scale for $\bm{\SphyNeu}$ mutations.} Each stack bar corresponds to the DFE of new mutations predicted to be $\SphyNeu$ by the mutation selection model, in a given population. The dashed area represents the proportion of deleterious mutations ($\SpopDel$), the gray area that of neutral mutations ($\SpopNeu$), and the yellow area that of beneficial mutations ($\SpopBen$).\label{fig:pop-weak}}
    \end{center}

    \begin{center}
        \begin{minipage}{0.75\linewidth}
            \begin{minipage}{0.9\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/Theta.pos.stacked.pdf}
            \end{minipage}
            \begin{minipage}{0.09\linewidth}
                \includegraphics[width=\linewidth, page=1]{artworks/legend.polycat}
            \end{minipage}
        \end{minipage}
    \captionof{figure}[Estimation of selection at the population scale for $\SphyBen$ mutations.]{\textbf{Estimation of selection at the population scale for $\bm{\SphyBen}$ mutations.} Each stack bar corresponds to the DFE of new mutations predicted to be $\SphyBen$ by the mutation selection model, in a given population. The dashed area represents the proportion of deleterious mutations ($\SpopDel$), the gray area that of neutral mutations ($\SpopNeu$), and the yellow area that of beneficial mutations ($\SpopBen$).\label{fig:pop-pos}}
    \end{center}

    \newpage
    \subsection{Distribution of scaled selection coefficients (\texorpdfstring{$\Sphy$}{S₀})}\label{subsec:expectedDFE}

    Theoretically, the selection coefficient ($\Sphy$) for any mutation is related to its probability of fixation by the equation $\proba_{\text{fix}} (\Sphy) = \Sphy/(1-\exp(-\Sphy)) $.
    It is thus possible to predict the distribution of $\Sphy$ of substitutions given the distribution of $\Sphy$ for mutations (Figure~\ref{dfe-sfs}A for one population of \textit{Chlorocebus sabaeus}) by weighting the different mutations by their $\proba_{\text{fix}} (\Sphy)$.
    The predicted DFE (distribution of fitness effects) of substitutions (Figure~\ref{dfe-sfs}B) matches the observed DFE of substitutions (Figure~\ref{dfe-sfs}C) in the terminal lineage, meaning that the $\Sphy$ obtained by phylogenetic mutation-selection codon models is a strong predictor of selection effectively exerted on mutations in terminal lineages.
    Equivalent figures to Figure~\ref{dfe-sfs} for each one of the 28 populations are available at Zenodo (\url{https://doi.org/10.5281/zenodo.7878953}) in file \textit{distribution\_S0\_28\_populations.pdf}.\\

    \begin{center}
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny A: Expected DFE for mutations}
            \includegraphics[width=\linewidth, page=1]{artworks/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.pdf}
        \end{minipage}
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny B: Expected DFE for substitutions}
            \includegraphics[width=\linewidth, page=1]{artworks/DFE.Chlorocebus_sabaeus.Saint_Kitts.MutSel.Flow.pdf}
        \end{minipage}
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny C: Observed DFE for substitutions}
            \includegraphics[width=\linewidth, page=1]{artworks/subs.Chlorocebus_sabaeus.Saint_Kitts.bins3Mask0.9.pdf}
        \end{minipage}
        \\
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny D: Observed DFE for polymorphisms}
            \includegraphics[width=\linewidth, page=1]{artworks/poly.Chlorocebus_sabaeus.Saint_Kitts.bins3Mask0.9.pdf}
        \end{minipage}
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny E: Site frequency spectrum}
            \includegraphics[width=\linewidth, page=1]{artworks/Chlorocebus_sabaeus.Saint_Kitts.MutSel-sfs.normalize.pdf}
        \end{minipage}
        \begin{minipage}{0.32\linewidth}
            \flushleft {\tiny F: $S$ as function of $S_0$ for each class}
            \includegraphics[width=\linewidth, page=1]{artworks/Chlorocebus_sabaeus.Saint_Kitts.MutSel.polyDFE_C.pdf}
        \end{minipage}
        \captionof{figure}[Distribution of scaled selection coefficients ($\Sphy$).]{\textbf{Distribution of scaled selection coefficients ($\bm{\Sphy}$) in the Saint Kitts population of \textit{Chlorocebus sabaeus}.}
        A) Distribution of scaled selection coefficients ($\Sphy$), predicted for all possible non-synonymous DNA mutations away from the ancestral exome.
        Mutations are divided into three classes of selection: deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial non-adaptive ($\SphyBen$).
        B) Expected distribution of $\Sphy$ for substitutions obtained by transforming the distribution of $\Sphy$ for mutations (panel A) by weighting each mutation by its scaled probability of fixation given by $\proba_{\text{fix}} (\Sphy) = \Sphy/(1-\exp(-\Sphy))$.
        C) Distribution of scaled selection coefficients ($\Sphy$) for all observed substitutions along the terminal branch of the phylogenetic tree.
        D) Distribution of scaled selection coefficients ($\Sphy$) for all observed polymorphisms currently segregating in the population.
        E) The site-frequency spectrum (SFS) in the population for a random sample of 16 alleles (means in solid lines and standard deviations in color shades) for each class of selection and for synonymous mutations, supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis). At high frequencies, deleterious mutations are underrepresented.
        F) Proportion of beneficial $\ProbaPopDel$, nearly-neutral $\ProbaPopNeu$, and deleterious mutations $\ProbaPopBen$  estimated at the population scale for each class of selection at the phylogenetic scale. Proportions depicted here are not weighted by their mutational opportunities.}
        \label{dfe-sfs}
    \end{center}

    \newpage
    \section{Correlation with diversity}

    \subsection{Phylogenetic Generalized Linear Model}

    Because a correlation must account for phylogenetic relationship and non-independence of samples, we fitted a Phylogenetic Generalized Linear Model (PGLM) in R with the package caper~\cite{orme_caper_2013}, with multi-furcation of the different populations inside each species.
    The mammalian tree imported from TimeTree~\cite{kumar_timetree_2017} and pruned to the species used in this study (Figure~\ref{tree}) is shown in newick format:

    \lstinputlisting[breaklines]{artworks/PGLM.tree}

    \begin{center}
        \includegraphics[width=0.65\linewidth, page=1]{artworks/PGLM.pdf}
        \captionof{figure}[Tree used for PGLM (Phylogenetic Generalized Linear Model).]{\textbf{Tree used for PGLM (Phylogenetic Generalized Linear Model).} The mammalian dated tree was obtained from TimeTree~\cite{kumar_timetree_2017} and pruned to include only the species analysed in this study, with multi-furcation of the different populations from each species placed at the same divergence time as the divergence to the sister species.}
        \label{tree}
    \end{center}

    \newpage
    Accounting for phylogenetic relationship and non-independence of samples, through the fit of a Phylogenetic Generalized Linear Model, we can estimate regression coefficient between two variables.
    Thus, for each population, the proportion of deleterious ($\ProbaPopDel$,~\ref{fig:pop-neg-prop}), nearly-neutral ($\ProbaPopNeu$,~\ref{fig:pop-weak-prop}) and beneficial ($\ProbaPopBen$,~\ref{fig:pop-pos-prop}) mutations estimated at the population-genetic scale is shown as function of $\Ne$, and we estimated regression coefficients.

    \subsubsection{Proportion of deleterious mutations (\texorpdfstring{$\SpopDel$}{D})}\label{subsec:proportion-deleterious-mutations}

    Higher effective population size ($\Ne$) is typically accompanied by a higher proportion of effectively deleterious mutations at the population scale ($\proba{[} \SpopDel {]}$), as shown in~\ref{fig:pop-neg-prop} (left panel).
    This trend is also confirmed when we restricted the analysis to class of mutations that are supposedly deleterious at the phylogenetic scale ($\SphyDel$), as shown in~\ref{fig:pop-neg-prop} (right panel).
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that very slightly deleterious mutations are more efficiently purified in large populations.

    \begin{center}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.all_P-Sneg.scatter.pdf}
    \end{minipage}
    \begin{minipage}{0.49\linewidth}
        \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.neg_P-Sneg.scatter.pdf}
    \end{minipage}
    \captionof{figure}[Relationship between the proportion of deleterious mutations ($\SpopDel$) and effective population size.]{\textbf{Relationship between the proportion of deleterious mutations ($\bm{\SpopDel}$) and effective population size.}
    Left) Proportion of deleterious mutations at the population scale ($\proba{[} \SpopNeu{]}$), shown as a function of estimated effective population size ($\Ne$). Right) Probability for a mutation to be deleterious at the population scale, given it is predicted to be deleterious at the phylogenetic scale, as a function of estimated effective population size ($\Ne$). Populations are represented by circles, and squares are the mean for the species. Correlations account for phylogenetic relationship and non-independence of samples, through the fit of a Phylogenetic Generalized Linear Model (see Materials \& Methods).\label{fig:pop-neg-prop}}
    \end{center}

    \subsubsection{Proportion of nearly-neutral mutations (\texorpdfstring{$\SpopNeu$}{N})}\label{subsec:proportion-nearly-neutral-mutations}
    Higher effective population size ($\Ne$) is typically accompanied by a smaller proportion of neutral mutations at the population scale ($\proba{[} \SpopNeu {]}$ in the range 0.06-0.18), as shown in~\ref{fig:pop-weak-prop} (left panel).
    This result is more pronounced ($\proba{[} \SpopNeu \given \SphyNeu {]}$ in the range 0.36-0.73) when we restricted the analysis to the class of mutations that are supposedly nearly-neutral at the phylogenetic scale ($\SphyNeu$), as shown in~\ref{fig:pop-pos-prop} (right panel).
    This result suggests that in populations with higher diversity (e.g.,~\textit{Bos} or \textit{Ovis}), mutations are more likely to be detected as under selection (positive or negative).
    Alternatively stated, mutations in populations with low diversity (e.g.,~\textit{Homo}) are effectively nearly-neutral and behave as neutral mutations would.
    This result is qualitatively in accordance with the nearly-neutral theory of evolution which argues that mutations are less efficiently selected for in small populations.

    \newpage
    \begin{center}
        \begin{minipage}{0.49\linewidth}
            \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.all_P-Sweak.scatter.pdf}
        \end{minipage}
        \begin{minipage}{0.49\linewidth}
            \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.weak_P-Sweak.scatter.pdf}
        \end{minipage}
        \captionof{figure}[Relationship between the proportion of nearly-neutral mutations ($\SpopNeu$) and effective population size.]{\textbf{Relationship between the proportion of nearly-neutral mutations ($\bm{\SpopNeu}$) and effective population size..}
        Left) Proportion of deleterious mutations at the population scale ($\proba{[} \SpopNeu{]}$), shown as a function of estimated effective population size ($\Ne$). Right) Probability for a mutation to be nearly-neutral at the population scale, given it is predicted to be nearly-neutral at the phylogenetic scale, as a function of estimated effective population size ($\Ne$). Populations are represented by circles, and squares are the mean for the species. Correlations account for phylogenetic relationship and non-independence of samples, through the fit of a Phylogenetic Generalized Linear Model (see Materials \& Methods).\label{fig:pop-weak-prop}}
    \end{center}

    \subsubsection{Proportion of beneficial mutations (\texorpdfstring{$\SpopBen$}{B})}\label{subsec:proportion-beneficial-mutations}
    A higher effective population size ($\Ne$) is accompanied by a smaller proportion of beneficial mutations at the population scale ($\proba{[} \SpopBen {]}$), as shown in~\ref{fig:pop-pos-prop} (left panel).
    This trend is also confirmed when we restricted the analysis to the class of mutations that are supposedly non-adaptive beneficial at the phylogenetic scale ($\SphyBen$), as shown in~\ref{fig:pop-pos-prop} (right panel).

    \begin{center}
        \begin{minipage}{0.49\linewidth}
            \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.all_P-Spos.scatter.pdf}
        \end{minipage}
        \begin{minipage}{0.49\linewidth}
            \includegraphics[width=\linewidth, page=1]{artworks/results.pop_size.pos_P-Spos.scatter.pdf}
        \end{minipage}
        \captionof{figure}[Relationship between the proportion of beneficial mutations ($\SpopBen$) and effective population size.]{\textbf{Relationship between the proportion of beneficial mutations ($\bm{\SpopBen}$) and effective population size.}
        Left) Proportion of beneficial mutations at the population scale ($\proba{[} \SpopNeu{]}$), shown as a function of estimated effective population size ($\Ne$). Right) Probability for a mutation to be beneficial at the population scale, given it is predicted to be beneficial at the phylogenetic scale, as a function of estimated effective population size ($\Ne$). Populations are represented by circles, and squares are the mean for the species. Correlations account for phylogenetic relationship and non-independence of samples, through the fit of a Phylogenetic Generalized Linear Model (see Materials \& Methods).\label{fig:pop-pos-prop}}
    \end{center}

    \section{Probability for beneficial mutations to be non-adaptive (\texorpdfstring{$\proba{[}\SphyBen\given \SpopBen {]}$}{P[B₀|B]})}

    \subsection{Across different populations}
    Probability for beneficial mutations to be non-adaptive, namely $\proba{[}\SphyBen\given \SpopBen {]}$, is obtained from Bayes' formula (eq.~14) merging three terms: the probability for a new mutation to be a non-adaptive beneficial mutation ($\proba{[} \SphyBen {]}$, eq.~5), the probability for a mutation to be beneficial ($\proba{[} \SpopBen {]}$, eq.~15) and the probability for a mutation to be beneficial at the population scale given that it is predicted to be a non-adaptive beneficial mutation at the phylogenetic scale ($\proba{[} \SpopBen \given \SphyBen{]}$, eq.~12).
    These four estimates across populations are shown in Table~\ref{table:proba}.
        \begin{longtable*}{|l|l|r|r|r|r|r|r|}
            Population & Species & $N_e$ & $\proba{[}\SphyBen{]}$ & $\proba{[} \SpopBen {]}$ & $\frac{\proba{[}\SphyBen]}{\proba{[} \SpopBen ]}$ & $\proba{[} \SpopBen \given \SphyBen{]}$ & $\proba{[}\SphyBen\given \SpopBen {]}$ \\
            \multicolumn{8}{r}{{Continued on next page}} \\
            \rowcolor{LIGHTGREY} Equus c. & Equus caballus & $7.5\times 10^{4}$ & $ 0.012$ & $ 0.015$ & $ 0.827$ & $ 0.648$ & $ 0.536$ \\
            Iran & Bos taurus & $5.6\times 10^{4}$ & $ 0.011$ & $ 0.039$ & $ 0.279$ & $ 0.873$ & $ 0.243$ \\
            Uganda & Bos taurus & $1.3\times 10^{5}$ & $ 0.011$ & $ 0.015$ & $ 0.720$ & $ 0.576$ & $ 0.415$ \\
            \rowcolor{LIGHTGREY} Australia & Capra hircus & $1.7\times 10^{5}$ & $ 0.011$ & $ 0.023$ & $ 0.480$ & $ 0.368$ & $ 0.177$ \\
            \rowcolor{LIGHTGREY} France & Capra hircus & $1.9\times 10^{5}$ & $ 0.011$ & $ 0.022$ & $ 0.515$ & $ 0.368$ & $ 0.190$ \\
            \rowcolor{LIGHTGREY} Iran (C. aegagrus) & Capra hircus & $1.9\times 10^{5}$ & $ 0.011$ & $ 0.025$ & $ 0.448$ & $ 0.368$ & $ 0.165$ \\
            \rowcolor{LIGHTGREY} Iran & Capra hircus & $2.3\times 10^{5}$ & $ 0.011$ & $ 0.021$ & $ 0.525$ & $ 0.368$ & $ 0.193$ \\
            \rowcolor{LIGHTGREY} Italy & Capra hircus & $1.9\times 10^{5}$ & $ 0.011$ & $ 0.017$ & $ 0.660$ & $ 0.368$ & $ 0.243$ \\
            \rowcolor{LIGHTGREY} Morocco & Capra hircus & $2.2\times 10^{5}$ & $ 0.011$ & $ 0.017$ & $ 0.667$ & $ 0.368$ & $ 0.245$ \\
            Iran & Ovis aries & $3.8\times 10^{5}$ & $ 0.012$ & $ 0.006$ & $ 1.984$ & $ 0.205$ & $ 0.407$ \\
            Iran (O. orientalis) & Ovis aries & $4.5\times 10^{5}$ & $ 0.011$ & $ 0.012$ & $ 0.983$ & $ 0.193$ & $ 0.190$ \\
            Iran (O. vignei) & Ovis aries & $3.7\times 10^{5}$ & $ 0.012$ & $ 0.020$ & $ 0.579$ & $ 0.190$ & $ 0.110$ \\
            Various & Ovis aries & $4.1\times 10^{5}$ & $ 0.011$ & $ 0.012$ & $ 0.967$ & $ 0.229$ & $ 0.222$ \\
            Morocco & Ovis aries & $ 4\times 10^{5}$ & $ 0.012$ & $ 0.005$ & $ 2.435$ & $ 0.211$ & $ 0.514$ \\
            \rowcolor{LIGHTGREY} Barbados & Chlorocebus sabaeus & $1.1\times 10^{5}$ & $ 0.009$ & $ 0.021$ & $ 0.452$ & $ 0.648$ & $ 0.293$ \\
            \rowcolor{LIGHTGREY} Central Afr. Rep. & Chlorocebus sabaeus & $1.7\times 10^{5}$ & $ 0.009$ & $ 0.018$ & $ 0.515$ & $ 0.535$ & $ 0.275$ \\
            \rowcolor{LIGHTGREY} Ethiopia & Chlorocebus sabaeus & $1.4\times 10^{5}$ & $ 0.009$ & $ 0.021$ & $ 0.444$ & $ 0.552$ & $ 0.245$ \\

>>>>

\begin{document}
    \maketitle
    \part*{S2 File}
    \vspace{-1em}
    \tableofcontents
    \newpage
    \listoffigures
    \listoftables
    \newpage
    \section{Beneficial mutations in the terminal lineages and populations}\label{sec:beneficial-mutations}

@@
